### 2010-03-02 paper1/hotpatch/abstract.tex
added paper to git repo (try 2)

\begin{abstract}
Despite advances in software modularity, security, and reliability,
offline patching remains the predominant form of updating or
protecting commodity software.  Unfortunately, the mechanics of {\it
  hot patching} (the process of upgrading a program while it executes)
remain understudied, even though such a capability offers practical
benefits for both consumer and mission-critical systems.
%
A reliable hot patching procedure would serve particularly well by
reducing the downtime necessary for critical functionality or security
upgrades.  Yet, hot patching also carries the risk -- real or
perceived -- of leaving the system in an inconsistent state, which
leads many owners to forego its benefits as too risky.  
%
%MEL:
%entity -- an
%application binary, library or even the kernel, while it is still in
%execution. 
%SB:
% It offers enormous benefits as it helps in mitigating
% process downtime and contributes to overall system reliability and
% availability.
%
%It offers great practical benefits for mission-critical systems, by
%reducing downtime for upgrades essential for maintaining their {\em
%  reliability} and {\em availability}. 
%
In this paper, we propose a novel method for hot patching ELF binaries
that supports {\em (a)} synchronized global data and code updates and
{\em (b)} reasoning about the results of applying the hot patch.  We
propose a format, which we call a {\it Patch Object}, for encoding patches as
a special type of ELF relocatable object file.  Our tool, {\em
  Katana}, automatically creates these patch objects as a by-product of the
standard source build process.  Katana also allows an end-user to
apply the Patch Objects to a running process.  In essence, our method
can be viewed as an extension of the Application Binary Interface
({\it ABI}), and we argue for its inclusion in future ABI standards.

%MEL: tried to condense to above
%We represent our hot patches as ELF-based Patch Objects, and we rely on the
%standard ELF structures in the running process' image for their
%application. 
%These Patch Objects are a special type of ELF relocatable object files,
%and are created automatically as a by-product of the standard source 
%build process by the developer-side component of our tool {\em Katana}, 
%and can then be applied by Katana's end-user binary-only component.  
%

%
%SB: This is ``how'', and is best relocated to the Intro.
% by first recognizing the importance of the stereotypical
% build process in constructing a binary through intermediate object
% files, and then delegating the overhead of maintaining object file
% dependencies to this build process. 
% In effect, what we are saying is
% that, after making the necessary source-level changes for a patch or
% an upgrade, we observe the changes to all object files and only
% consider those that have been modified or created anew by the build
% process. Correlating these with the original object files that were
% used to build the previous version of the executable, we come up with
% a mechanism to upgrade the running process with the new binary.
\end{abstract}