### 2010-03-02 paper1/hotpatch/acks.tex

% superceded by \thanks in header
\section*{Acknowledgements}

{\small 
The first three authors' work was supported in part by the National
Science Foundation, under grant CNS-0524695.  The views and
conclusions do not necessarily represent those of the sponsors.
%
Locasto is supported in part by grant
2006-CS-001-000001 from the U.S. Department of Homeland Security under
the auspices of the I3P research program. The I3P is managed by
Dartmouth College. The opinions expressed in this paper should not be
taken as the view of the authors' institutions, the DHS, or the I3P.
}