### 2010-03-02 paper1/hotpatch/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

We introduce a method for hot patching: a technique we believe to be a
promising alternative to redundancy, ad hoc self--healing techniques,
``patch and pray,'' or other approaches to dynamic software updates.
Hot patching has the potential for aligning actual practices with
acknowledged ``best practices'' relating to critical security or 
functionality updates.  We hold that one major impediment to hot
patching is the opaque nature of most patches (be it
proprietary or open software), and our method of patching along with the
PO file format are first attempts at providing a basis for informed reasoning
about the structure and implications of a patch.