### 2010-03-02 paper1/hotpatch/design.tex

\section{Katana Design: Tracking Object Dependencies}
\label{sec:design}

Katana leverages the typical Unix {\tt Makefile} build mechanism to track
file-level dependencies.  Normally, after applying a source patch file
and performing a top-level {\tt make}, only those object files whose
underlying sources have changed are rebuilt.  Katana thus tracks
object-level (.o) dependencies as follows.  We first replicate all
object files and the ELF executable from the existing source tree.  We
then apply the patch to the {\it original} source tree.  At this
point, only source files have been modified. Next, using the Linux
kernel's {\it inotify} mechanism\footnote{{\it inotify} allows the
  registration of filesystem triggers. We use inotify to avoid potentially
  expensive recursive directory comparisons.}, Katana sets up a notification
on the original source tree, so that it knows when an object file is
created or modified under the original tree.  Finally, we perform the
top-level {\tt make} under this source and record all created/modified
object files, along with the newly created executable. The use of
inotify is not inherent to our general patching mechanism, and indeed
our present prototype only compares two directory trees and examines
all object files.

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.25]{making.pdf}
\end{center}
\caption{{\small {\it An Example Code Base.} From the top: each source
    file creates a corresponding object file; multiple object files
    are combined into intermediate compilation units (CU); and
    multiple CUs are merged to form the executable. All shaded blocks
    indicate modified files.}}
\label{fig:making}
\end{figure}

Figure~\ref{fig:making} illustrates how object files are modified by
{\tt make} as a result of source-level changes.  Katana only considers
objects that are closest to the source files and ignores all other
intermediate object files and compilation units (CU).  Hence, in
Figure~\ref{fig:making}, Katana only records the shaded circle-objects
along with the final executable.

To dynamically update the running application, Katana needs to patch
both the {\bf code} and the {\bf data} within the process.  It first
creates a {\it patch object} (PO): an ELF file with sections that
indicate the type of patch (code or data), the patch offsets and
lengths within the process address space, patch data, function and
data names, {\it etc.} 
%The following sections explain how Katana
%handles both code and data patching.