### 2010-03-02 paper1/hotpatch/main.tex

%\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage{usenix}
%\usepackage[margin=1in]{geometry}

\documentclass{IEEEtran}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{authblk}

\title{Katana: A Hot Patching Framework for ELF Executables}

\author[1]{Ashwin Ramaswamy}
\author[1]{Sergey Bratus}
\author[2]{Michael E.\ Locasto}
\author[1]{Sean W. Smith}
\affil[1]{Dartmouth College}
\affil[2]{George Mason University}

\begin{document}

\maketitle
\pagestyle{plain}

\input{abstract}

{\footnotesize
\begin{quote}
``Some reports, such as the case of the Conficker outbreak within 
Sheffield Hospital's operating ward, suggest that even security-
conscious environments may elect to forgo automated software patching, 
choosing to trade off vulnerability exposure for some perceived notion 
of platform stability...'' -- \url{http://mtc.sri.com/Conficker/}
\end{quote}
}

\input{intro}

\input{challenges}

\input{redundancy}

\input{design}

\input{patching}

%\input{discussion}

\input{poformat}

\input{related}

\input{conclusion}

{\small
\bibliographystyle{abbrv}
\bibliography{locasto,extras,oakley}
}

\end{document}