### 2010-03-02 paper1/hotpatch/related.tex

\section{Related Work}
\label{sec:related}

There are several hot-patching systems preceding Katana. One of the
most well known is probably Ginseng \cite{ginseng}. Ginseng (and
systems drawing inspiration from it such as Polus \cite{polus}) have
successfully demonstrated patching of such important software as
apache and sshd. These systems perform analysis of the differences
between the original and patched versions at the source code
level. This introduces considerable (and we argue unnecessary)
complexity and inability to deal well with some optimizations such as
inlining and hand-written assembly. The complexity of analyzing the
source code ties these systems to generally a single language (C in
the case of both Ginseng and Polus). By contrast, Katana is language
agnostic as it works at the level of the binary ABI, and although we
have not yet demonstrated it doing so, it should eventually be able to
patch binaries compiled from any language, providing the necessary
symbol and relocation information is supplied. Ginseng also requires
significant programmer interaction in annotating the code
\cite{ginseng-manual} and requires compiling the code to use
type-wrappers, allowing the patching of data types but at the cost of
indirect access to them. The more programmer effort involved in
generating a patch, the more likely the patch is to be incomplete or
incorrect.

Motivated by many of the points in the above paragraph, the
successful Ksplice system \cite{ksplice} patches at the binary level, as we do. We
claim the following differences from and improvements over Ksplice.
\begin{itemize}
\item Ksplice operates on the kernel. As their paper states, most of their
  technique is not specific to the kernel, but there is no evidence
  that it has been implemented to function on userland programs.
  Katana operates on userland.
\item Ksplice makes no attempt to patch data, relying entirely on
  programmer-written transformation functions when data types do
  change
\item Ksplice patches are created as kernel modules. Ksplice does not
  provide a mechanism to perform operations, such as composition,
  on these patches.
\end{itemize}

To the best of our knowledge, Katana is the first system to utilize
DWARF type information in patching.

Maintaining continuous availability, even in the absence of disruptive
events like patches, is both a challenging technical exercise and the
driving need for research on dependability, reliability, and fault
tolereance~\cite{deltaexec}. 
%The work most closely related to Katana focuses
%JMHO: reworded above because I don't think this is what's most
%closely related
Our work follows work focusing
%JMHO end
 on enabling a software
application to continue providing service or survive significant
events like errors, exploits, and patches.  This body of work includes
research on dynamic kernel updates, software survivability, and
software self-healing. However, other research areas also addressed the
challenge of enabling software to adapt at runtime, e.g., the area
of software evolution (e.g.,~\cite{software-evolution}).

The concept of crash-only software~\cite{crashonly} advocates
microrebooting: the procedure of retrofitting each component of a
system with the ability to crash and reboot safely as the default mode
of operation.  Despite its appeal as a design principle, such an
approach would be difficult to retrofit to legacy software.
Although restarting a particular service or application is disruptive
enough, rebooting the operating system itself multiplies this
disruption.  
The need to avoid that kind of downtime helped drive the
creation of frameworks like Loadable Kernel Modules for Linux, which
allow for extending the kernel during runtime without a reboot.
The ability to update the running kernel (as opposed to adding
or removing modules) without rebooting was achieved at least ten years
ago~\cite{cesare1998} and recently rediscovered, albiet mostly for
research, rather than commodity,
kernels~\cite{nonLKMkernpatch,baumann2007osupdate,soules2003k42}).
Even so,
dynamic updates of the kernel during runtime that don't require a
reboot are difficult to apply to a commodity OS, although several
efforts have been successful for the K42 experimental
system~\cite{soules2003k42,baumann2005k42}.

Software self-healing aims at ensuring continuous or increased
availability for systems subjected to exploited vulnerabilities,
either by automatically generating
patches~\cite{weimer2009genetic,stelios2005usenix} to gradually harden
the application or seeking to avoid a restart altogether by modifying
certain runtime aspects ({\it e.g.,} the memory
subsystem~\cite{rinard2004osdi}, properties of the execution
environment~\cite{rx2005sosp}), 
%
or selected control
paths~\cite{dira2005,locasto2007usenix}) of the system in response to
attacks.
One major risk of employing self-healing in production environments is
that the semantics of follow-on execution remain largely uncontrolled,
although recent work in automatically correcting memory
errors~\cite{exterminator2008cacm} seems to achieve fairly reliable
results.  Both automated responses and traditional patches can make it
difficult for an administrator to understand the implications of a
particular fix~\cite{rinard.ppp}.