### 2010-03-02 paper1/hotpatch/secse-main.tex

\documentclass[10pt, conference, compsocconf]{IEEEtran}
\IEEEoverridecommandlockouts % for \thanks
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{url}
%\usepackage{hyperref} -- DISABLED: fails ``PDF express'' check

% \usepackage{authblk} -- must use IEEE's own
% \usepackage[margin=1in]{geometry}

\title{Katana: A Hot Patching Framework for ELF Executables}

\author{\IEEEauthorblockN{Sergey Bratus, James Oakley, Ashwin Ramaswamy, 
Sean W.\ Smith\thanks{The first four authors' work was supported in part by the National
Science Foundation, under grant CNS-0524695.  The views and
conclusions do not necessarily represent those of the sponsors.}}
\IEEEauthorblockA{Computer Science Dept.\\
  Dartmouth College\\
  Hanover, New Hampshire}
\and
\IEEEauthorblockN{Michael E.\ Locasto\thanks{Locasto is supported in part by grant
2006-CS-001-000001 from the U.S. Department of Homeland Security under
the auspices of the I3P research program. The I3P is managed by
Dartmouth College. The opinions expressed in this paper should not be
taken as the view of the authors' institutions, the DHS, or the I3P.}}
\IEEEauthorblockA{Computer Science Dept.\\
George Mason University\\
Arlington, Virginia}}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\maketitle
%\pagestyle{plain}

\input{abstract}

{\footnotesize
\begin{quote}
``Some reports, such as the case of the Conficker outbreak within 
Sheffield Hospital's operating ward, suggest that even security-
conscious environments may elect to forgo automated software patching, 
choosing to trade off vulnerability exposure for some perceived notion 
of platform stability...'' -- \url{http://mtc.sri.com/Conficker/}
\end{quote}
}

\input{intro}

\input{challenges}

\input{likelinking}

\input{redundancy}

\input{design}

\input{patching}

\input{discussion}

\input{poformat}

\input{related}

\input{conclusion}

% Now in \thanks in author list.
%\input{acks}

{\small
\bibliographystyle{abbrv}
\bibliography{locasto,extras,oakley}
}

\end{document}