### 2010-03-18 paper1/hotpatch/response_to_reviews.txt
added response to reviewers

The following document details how we handled each comment for each of
the reviewers for IJSSE paper #2493. Comments from the reviewers are
quoted using the characters >> at the start of each line.

REVIEW 1
>> I urge the authors to complete the missing work, especially in
>> the area of safety and multithreading. Both are prerequisites for
>> this work being acceptable in the real world.
We have actually made significant work on implementing safety analysis
since the submission for review, so we updated the paper slightly with
that information. We added a slightly greater nod to the importance of
multithreading (and our intent to address it) in the Challenges
subsection of section III.

>> Section I should mention virtualization as a path to upgrading
>> existing systems with patches, basically rolling upgrades. While
>> this won't work for monolithic systems, it does work extremly well
>> for horizontally scaled software architectures and eliminates the
>> need for hotpatching. If traditional patching is - as you claim -
>> often not done due to the preception of instability, why would
>> hotpatching be perceived as a better choice?
Virtualization is a type of redundancy. The hardware expense and waste
does not apply to it, but our arguments about state loss and
complexity do apply. We have clarified our argument about redundancy
slightly to show that this case is included.

>> I would redraw Fig.1 as follows: - remove second patcher - add arrow
>> from patcher back into runtime. Basically, the future/patched runtime
>> should not appear any different that the current unpatched
>> runtime. Think about it, in the next phase the previously patched
>> runtime is the basis you are starting from. - Rename "Patched Runtime"
>> to something like "Hotpatch application phase", etc.
We disagree. We are showing the lifecycle of software and how it is
extended, not that it is running differently after it is patched. The
point about the name of that stage is well taken, however, and it has
been renamed.

>> Your description includes only sequential patch application (n-1 ->
>> n). What about wanting to upgrade from n-5 to n?  
This is entirely possible, and we have added a slight clarification in
section V to make this harder to miss

REVIEW 2 
>> The quote after the abstract is not necessary. If the authors
>> strongly feel about it, they can include that in an appropriate place
>> of the main text of the paper.
The quote has been moved to a more appropriate location.

>> The extension from the workshop paper is not clear, it may be
>> mentioned briefly by referring to the workshop paper in the
>> Introduction or the related works.

>> The title of both workshop and journal papers same. Based on the new
>> addition in the journal version, the title of the paper needs to be
>> changed.
The title has been changed

>> Section VII could be expanded with the details of major contributions,
>> lessons learned, limitations and future work.
Section VII has been expanded slightly. Much of this is covered
elsewhere in the paper, so to repeat it would be to risk excessive
redundancy.

REVIEW 3
>> The quote concerning conficker at the start of the paper is an
>> interesting anecdote but should perhaps be placed where you make the
>> argument that some trade security for perceived stability. 
The quote has been moved to a more appropriate location.

>> Your argument concerning redundancy is a bit of a strawman. Redundancy
>> is primarily used to ensure high availability, and in that context the
>> cost and resource is can be motivated. The ability to upgrade systems
>> with minimal disruption is just a nice byproduct. For patching
>> purposes alone, one could operate with a minimal amount of extra
>> resources (negating the "expensive" and "wasteful" arguments). Complex
>> and specialized do still apply in many (but far from all) cases.
We disagree. We are not arguing that redundancy is useless, merely
that it is not as attractive as a ubiquitous patching
mechanism. Building a system with redundancy advanced enough that
state can be transferred from one instance of a task to another
without requiring per-application coding of the state transferral is
not known to us to be a solved problem, and the reviewer points us to
no evidence that it is.

>> Figure one is far too large. 
This has been remedied

>> First you say Katana uses inotify, then that your current prototype
>> doesn't use it. Which is it? 
The original plan was to use inotify. We have never yet implemented its
use, as it is not central to patching and other features were more
important. We have therefore modified the section mentioning it to
give it only a very brief mention as "future work".

>> Data patching is the most challenging part of the problem. You discuss
>> this, but it isn't clear exacly what the limits of your approach are.
>> Can you handle pointers to members of structures (when the structure
>> changes)? Can you handle arrays? Pointers to array elements? Changed
>> meaning of data (e.g. if an element is added to an enum)?
>> Variable-size structures (i.e. those with a variable-size buffer at
>> the end)? Will pointer casts cause problems (e.g. if a pointer to a
>> changed data object is stored in a void* or caddr_t variable)? Do you
>> handle multiple pointers to the same heap object (or the situation
>> where you have one pointer to the start of a heap object, and one to
>> an element inside it)? Do unions cause problems (e.g. union between an
>> int and a pointer)?  
These are very valid questions. Section III
has been expanded slightly in an effort to address these concerns and
the current limitations of our system.

>> The limitation to code that does not call previously unused functions
>> in dynamically loaded libraries seems quite serious. Since you have
>> chosen not to support this, presumably you do not consider it a severe
>> limitation of Katana. Can you explain why? 
This is very true. When we were designing the system, it was deemed
difficult. Given its importance, however, we have made plans to
support it. We have fixed this in the paper, and cite the technique
that will be used.

>> Katana will only update when the program is in a safe state. In real
>> programs, how often will the program never enter a safe state? What
>> practical experience do you have with this? What practical experiences
>> do you have so far?
Unfortunately, we have little. We will certainly be publishing
real-world results when we have optained them.