### 2010-03-27 paper1/hotpatch/patching.tex

\section{Automated Patching}

In this section, we describe our data and code patching methods.  We
note that, compared to previous work, our PO data structures
allow reasoning about the scope, extent, and impact of the patch ({\it
 e.g.,} whether it affects particular subsystems within the process).

%\subsection{Code Patching}
{\bf Code Patching.} %\hskip 0.1in
This process involves several stages:
%\newline
\newline
(i) {\it Code Identification:} Katana first needs to identify the
section(s) of text that need to be modified within the running
process. To do this, we consider the list of all modified object files
from our tracking step and identify all functions (both static and
global) within these files 
%SB:
from their symbol table. 
%JMHO: removed talk of copying all functions
Functions that differ between the original
and modified versions of an object file are copied into the PO and
marked as code.
%\newline
\newline
(ii) {\it Symbol Resolution:} After identifying all functions that
require a patch, we need to resolve outstanding symbol references
within each function. Typically, symbol resolution for an application
happens at both the linking stage (called {\it static linking} when
the symbol is present within another object file or archive), and the
execution stage (or {\it dynamic linking}, when the symbol is present
within a shared library).  All code relocations are identified in the
ELF sections {\tt .rel.text} and {\tt .rela.text}, within the object
files and the final executable.  Each relocation entry contains, among
other information, the code offset that requires relocation and the
outstanding symbol that provides this fix-up.

For each relocation entry, Katana copies the corresponding symbol into
the PO. The actual value of the symbol is not necessary in the PO
unless the target to be patched will be stripped, as the value of the
symbol can be retrieved from the running target while performing the
patching. This is key in allowing executables that have already been
patched to be patched again.
%Katana uses the replicated executable (from
%{\it before} the patch) to figure out the address of the symbol.  If
%the symbol was provided by another object file, then the symbol table
%of the old executable contains this final address, and we update the
%PO accordingly,
%SB:
%with this address as the patch target. 
If the symbol was
dynamic ({\it i.e.,} present in a shared library such as {\tt libc}),
then the fixup value is the address of a corresponding entry in the
procedure linkage table ({\it PLT}) of the executable.  The PLT is
essentially a jump table with entries for each symbol that needs to be
resolved at runtime by the dynamic linker.  When the process begins
execution, the dynamic linker maps the required shared libraries into
the address space of the process and updates each PLT entry.

For dynamic symbols, Katana traverses the PLT entries of the
executable and compares the symbol name of each entry with
the symbol name that requires relocation. Once a match is found, the
symbol value can be determined. Our current prototype cannot add calls
to previously unused functions in shared libraries, but support for
this will be added using the ``ALTPLT'' technique described by \cite{elf-debugging}.

%JMHO: this got too many negative comments from reviewers, removing
%and clarifying above
% Katana is unable to handle calls to previously
% %SB:
% unused functions present in any shared libraries\footnote{This would
%   require creation of new PLT and GOT entries and either subsequent
%   re-basing of the following segments of the executable, or creation
%   of a new segment to allocate the extra entries. Although ELF
%   rewriting systems like ERESI or Diablo show that such manipulations
%   can be made practical, we chose not to complicate Katana with
%   them yet.}
% , although support for this is planned.

Finally, if the outstanding symbol's definition was not found within
the replicated executable (either within the symbol table or within the PLT),
then it was newly added by the patch; it is marked as such and added
to the PO.
%\newline
\newline
(iii) {\it Patch Application:} Applying a code patch is simple enough
and has been researched in other
systems~\cite{pannus,livepatch,ksplice,kaho}.  We map the new function
in memory and insert a trampoline {\tt jmp} instruction at the
beginning of the old function within the process memory image. This
interposition allows the caller to execute our new function instead of
the previous one at the cost of an extra jump.  It is possible to
avoid the overhead (from branch mis-prediction) of the {\tt jmp}
instruction by adding into the old function code that traces up the
stack and modifies the caller's {\tt call} instruction operand to
point to the new address instead of the old one. Although this
optimization would ensure that all subsequent calls from the same
caller would execute the new patched function without stepping into
the old one, it does makes the process of rolling back a patch
non-trivial. A simpler method to avoid the overhead would be to
relocate all calls to the function to point to its new definition
(although the trampoline would still be desirable, to catch calls from
function pointers or anything of similar nature). The current
prototype uses only the trampoline method.

%\subsection{Data Patching}
{\bf Data Patching.} %\hskip 0.1in
Patching data within a running process is significantly harder than
patching application code.  The primary challenge here is to
synchronize the code and the data structures it acts on.\footnote{For
  example, consider adding a new member to a C struct definition and
  an additional clause to the logic that processes it.}

Tracking down previously allocated data is nontrivial (one of the
reasons why garbage collectors are interwoven with the language
implementation).  Even after identifying the allocated chunks of
memory, in the absence of some kind of type specification, the {\it
  internal structure} of memory remains opaque.  We also need a method
for extracting only the modified data variables from the patch and a
means to discover the actual modifications that were performed. 
%JMHO: added the below because otherwise I don't think it's clear we
%solved them
Our system solves both of these problems.

We first note that any code that acts on patch-modified data is
already taken care of by Katana's code patching process
because we rely on {\tt make} to build the object files that
correspond to all modified sources. We resolve the previously
identified problems towards patching data by leveraging
DWARF\footnote{\url{http://dwarfstd.org}} debugging information within the
application executable. This requires the object files to be compiled
with debugging support, but we do not see this as a limitation. Since
we need DWARF information only while building the PO, all debugging
symbols could be stripped from the executable during application
deployment, if desired 
%JMHO:
(this would require storing symbol values in the PO, however).
%JMHO end
We recall the representation of types in
the DWARF format and then detail the various steps in Katana's data
patching process.

{\bf DWARF type information.} %\hskip 0.1in 
The DWARF structure is laid
out as a tree of DIEs ({\it Debugging Information Entries}) within the
executable file. Each DIE has an associated tag and a set of
attributes. The DIE that defines type information has the tag as one
of {\tt DW\_TAG\_base\_type}, {\tt DW\_TAG\_structure\_type} or {\tt
  DW\_TAG\_union\_type}. Typedefs and other type modifiers (such as
{\tt const}, {\tt volatile}, {\tt pointer} {\it etc.}) are referenced
by the DIE that defines the type. In case of structures or unions,
each member is contained as a separate DIE within the parent DIE that
identifies the struct/union. It is important to note that DWARF
annotates types of {\it all} visibilities from the program sources -
local, global and static.

Katana's data patching process contains a number of steps:

(i) {\it Type Discovery:} We set out to discover all newly created or
modified data types -- those that are primarily user-defined (such as
structures and unions in {\bf C}). Katana traverses the type information
(as identified by the above DWARF tags) from the newly created executable,
 and for each encountered type, it searches for
the corresponding type-name within the {\it replicated} executable
(from before the patch). If so found, the full types (i.e. the number,
type and position of all member variables contained within) are
compared to determine if they are identical. If not identical,
a transformation between the old and new versions is generated, along
with DWARF information identifying the type to insert into the PO as
soon as a variable is found making use of the altered type.
%the parent type identifying the struct/union is inserted into the PO.
 Else, if
the type name itself was not found within the replicated executable, then the
current type was created by the patch, and is added as such to the PO.
%\newline
\newline
(ii) {\it Data Traversal:} The next step is to traverse all variables
defined within the new application, and for each one encountered,
we first determine its 
%SB:
lexical scope.
If the scope is
local, then we ensure that the corresponding function (the one that
defines this variable) does not have an activation frame on  
the program stack while
applying the patch. Else, the variable has been defined as either
global or static. We first check whether the replicated executable defines
the same variable. If not, then this variable has been created by the
patch and we need not worry about it and may leave the symbol
resolution up to
the compiler (as only {\it new} code can use this variable). Otherwise,
we verify whether the variable's type is one of the
modified types identified during {\it type discovery}. If it is, then
we add the variable along with its {\it original} address from the replicated
executable, its {\it new} address from the patch, and its type information to the
PO. At the end of this stage, Katana would have identified all newly
created or modified variables from the patch. 
%\newline
\newline
(iii) {\it Patch Application: } Applying a data patch consists of
first tracking down the relevant symbols in program memory. Katana
reads in the PO, and for each data variable encountered, it checks whether
the variable is a pointer or not. If it is, then the current validity of the
pointer is verified (by bounds-checking the pointer value to within
heap boundaries).  If the pointer is found to be invalid, no further
action is taken.  If the pointer is valid, then memory for the new
type(s) is allocated, the older structure is copied into the new one
taking into account the difference in structure definition, the old
memory is then freed, and the pointer is modified to point to the new
segment (in case of structures such as lists, trees, since we have the
type specification, we can repeat this process recursively for each
node on the list or tree). Else if the variable is not a pointer, then
Katana modifies all its references in the program text to the updated
memory location from the patch. Katana supports default values in the
sense that if the variable has an initializer in the new version of
the program text, that initializer will be used for any member
variables within structures or array elements that did not
previously exist in the executing target. Initializers from the new
version are used for all \texttt{const} global variables.
Eventually Katana it will support default values and programmer-written
custom initializers.

{\bf Challenges.} %\hskip 0.1in 
Data patching, as described above, is not an easy task, even with
DWARF type information. C, as well as many other directly compiled
languages, does not have a strong type system, and is not designed to
allow reflection. Structures are relatively straightforward to patch,
because they contain a detailed specification. Unions and arrays, on
the other hand, are generally quite opaque. If substantial changes
have been made to one of the types in a union, Katana cannot do
anything automatically, as there is no automatic way to determine
which unioned type to act on. If an array changes size, Katana will
assume that it is growing or shrinking at the end and will copy old data
accordingly, but it will also issue a warning that this may not be the desired
behavior. Pointers are of course handled, but with some limitations
currently. Unfortunately for our purposes, memory management is not
part of the ABI. Further, there is no standard way to determine what
block a given pointer is part of. Our current implementation correctly
handles only memory management with \texttt{malloc} and only pointers to the
beginning of blocks. Improvement of this is a major area of future
work. Multiple pointers to the same address are handled by keeping track
of which addresses have been relocated. \texttt{void}* also poses a
problem, as there is generally no way to determine the ``real'' type
of the pointer. The general solution to all of these type problems is
to ask the programmer for routines which perform the
application-specific work. Minimizing programmer work is of foremost
importance, because greater human interaction adds greater possibility
for human error, thus possibly decreasing the reliability of the
patch, but it is not avoidable in all situations.

Hot patching still faces a number of other
challenges, including dealing with multithreaded programs and address
space randomization (which slight changes to the OS loader can help us
overcome).  
There is nothing inherent in our design which will not work with
multithreading, but dealing with it through ptrace takes considerable
work which we have not yet done. More importantly, the deadlock
avoidance work required to ensure that we do at some point pause all
threads but that we never pause a thread that another thread may be
waiting on in order to reach a safe state is nontrivial. We plan to
address this in the future.
% We consider the program to be in a safe state if all
% activation records are free of functions contained in the PO and all
% activition records are free of functions that $(1)$ access any global
% or static symbols we identify during Katana's {\it Data Traversal} stage and
% $(2)$ do not define any local variables of modified types identified
% during our Type Discovery phase.  Katana uses the {\tt ptrace}
% interface to pause execution and query this state.
%
%JMHO: We repeat this again in the discussion section, so I commented
%it out here and added a brief note saying we discuss it more below

% LocalWords:  ptrace multithreading JMHO