### 2010-04-02 paper1/hotpatch/secse-main.tex
formatting the way the publisher (IGI Global) wants it

\documentclass[12pt, conference, compsocconf]{IEEEtran}
%\documentclass[12pt]{article}
\IEEEoverridecommandlockouts % for \thanks
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{apalike}
\usepackage{endnotes}
\let\footnote=\endnote

%\sectionfont{\textbf}

%\usepackage{hyperref} -- DISABLED: fails ``PDF express'' check

% \usepackage{authblk} -- must use IEEE's own
% \usepackage[margin=1in]{geometry}

%\title{Katana: Towards Patching as a Part of the ELF Toolchain}
\title{Katana: Towards Patching as a Runtime part of the
  Compiler-Linker-Loader Toolchain}

\author{\IEEEauthorblockN{Sergey Bratus, James Oakley, Ashwin Ramaswamy, 
Sean W.\ Smith\thanks{The first four authors' work was supported in part by the National
Science Foundation, under grant CNS-0524695.  The views and
conclusions do not necessarily represent those of the sponsors.}}
\IEEEauthorblockA{Computer Science Dept.\\
  Dartmouth College\\
  Hanover, New Hampshire}
\and
\IEEEauthorblockN{Michael E.\ Locasto\thanks{Locasto is supported in part by grant
2006-CS-001-000001 from the U.S. Department of Homeland Security under
the auspices of the I3P research program. The I3P is managed by
Dartmouth College. The opinions expressed in this paper should not be
taken as the view of the authors' institutions, the DHS, or the I3P.}}
\IEEEauthorblockA{Computer Science Dept.\\
George Mason University\\
Arlington, Virginia}}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\maketitle
%\pagestyle{plain}

\addtoendnotes{ }%seems to be a bug in endnotes package, complains
                 %about missing .ent file if we don't add something

\input{abstract}

\noindent{\bf Keywords:} \input{keywords}

\input{intro}

\input{challenges}

\input{likelinking}

\input{redundancy}

\input{design}

\input{patching}

\input{discussion}

\input{poformat}

\input{related}

\input{conclusion}

% Now in \thanks in author list.
%\input{acks}

{\small
%\bibliographystyle{apa-good}
\bibliographystyle{apalike}
%\bibliographystyle{abbrv}
\bibliography{locasto,extras,oakley}
}

\newpage 
\parindent 0pt 
\parskip 2ex  
\def\enotesize{\normalsize}  
\theendnotes

%\theendnotes

\end{document}